Enforced Invariant Confluence

### 2020-08-08 doc/the_vldb_journal_paper/rebuttal.tex

\paragraph{Reviewer 4: 2}
\begin{feedback}
  Invariant confluence is about safety, however, could there be a liveness part
  of the problem worth investigating/discussing for replicated object
  management? Would it be possible that forgoing coordination across replicas
  could jeopardize/violate some liveness properties?
\end{feedback}
Forgoing coordination across replicas in favor of \invariantconfluence{}
actually \emph{increases} liveness, rather than jeopardizing or violating it.
Any alive replica of an \invariantconfluent{} object can service client
requests and execute transactions. This is the strongest form of liveness we
can hope to achieve. In fact, if we aim for strongly consistent replication,
then the CAP theorem tells us that total availability (the kind that
\invariantconfluence{} achieves) is impossible~\cite{brewer2012cap,
gilbert2002brewer}.

Note, however, that a segmented \invariantconfluent{} object does suffer from
liveness problems, as we discuss in Section 6.4. Specifically, transactions
that require global coordination are live only as long as every replica is
live. We discuss this more in Reviewer 4: 5 below.

\paragraph{Reviewer 4: 3}
\begin{feedback}
  Mergeable Replicated Data Types paper (OOPSLA'19) is relevant work. The paper
  does not cite MRDT understandably because it is very recent work, however
  including a discussion of this in the related work would be helpful.
\end{feedback}
\textcolor{red}{TODO.}

\paragraph{Reviewer 4: 4}
\begin{feedback}
  Section 4 is about invariant tightening, which has a broader applicability in
  distributed systems/programs. Unfortunately the techniques used here are not
  new. Involving the user for identifying unreachable states in not practical,
  and the discussion on tolerating user errors is weak as it does not offer a
  remedy. Would using some heuristics and abstracting state provide any help
  here?
\end{feedback}
We agree that involving the user for identifying unreachable states is not
ideal, and in many cases not practical. We think our discussion of tolerating
user errors is also not ideal, but not weak. By requiring users to provide
proofs of reachability and by deploying the invariant $I - NR$ instead of $I$,
invariant violations are impossible.
%
Heuristics and abstracting state and other techniques from invariant tightening
would probably help, but they are out of scope of the paper. As you mention,
the techniques are not new. We view it as an orthogonal but complementary line
of research.

### 2020-08-08 doc/the_vldb_journal_paper/rebuttal.tex

% Turn on the full reviewer comments if the turnonshowfeedback flag was passed
% in from the command line. This was taken from [1].
%
% [1]: https://stackoverflow.com/a/1466610/3187068
% \ifdefined\turnonshowfeedback
%   \toggletrue{showfeedback}
% \else
%   \togglefalse{showfeedback}
% \fi

\begin{document}
\begin{center}
  \ttlfnt Response to Reviewer Comments
\end{center}

\begin{multicols*}{2}
Below, you'll find our response to reviewer comments and our revisions. All
changes made to the paper are shown in \markrevisions{pink}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Reviewer 1}
\paragraph{Reviewer 1: W1}
\begin{feedback}
  More details about the DSL to specify the invariants with more complex
  examples would add a lot of value to the paper
\end{feedback}
We added a link to our DSL implementation on GitHub to the beginning of Section
8. The GitHub repository also contains a couple of examples that didn't make it
into the paper. We also added a bit more explanation of the DSL to Section 8.1.
We agree the paper would benefit from more examples, but given that the paper
already has over 10 examples of invariant confluence with varying degrees of
complexity from simple (e.g., Example 10) to complex (e.g, Example 14), we feel
the added benefit of yet another example does not justify the effort.

\paragraph{Reviewer 1: W2}
\begin{feedback}
  More experiments with some realistic workloads (like TPCC) to show the
  benefit of the proposed concepts in action are needed
\end{feedback}
See above. We agree that having a number of realistic workloads in the paper
would be beneficial, but specifying, verifying, and explaining every workload
requires a significant amount of work for a marginal increase in the clarity of
the paper.

\paragraph{Reviewer 1: W3}
\begin{feedback}
  Even though the automatic segmentation is left for future work, an example
  with a manual mapping of segments in a realistic workload (like TPCC) would
  show-case its potential
\end{feedback}
See above.

\paragraph{Reviewer 1: W4}
\begin{feedback}
  The main weak point of this specific submission is whether it has enough new
  material to to qualify for being an extended version
\end{feedback}
See D6.

\paragraph{Reviewer 1: D1}
\begin{feedback}
  W4 touches upon the ``new material'' for this extended version of the paper,
  and the above weak points W1, W2 and W3 can all serve as additional material.
\end{feedback}
See W1, W2, W3, and D6.

\paragraph{Reviewer 1: D2}
\begin{feedback}
  W1 will essentially help the reader to quickly see the value of the proposed
  definitions and theorems with some specification examples that will be able
  to capture more interesting use-cases.
\end{feedback}
See W1.

\paragraph{Reviewer 1: D3}
\begin{feedback}
  W2 would help to quantify the benefits that come from being able to have some
  invariant confluent objects in an otherwise strongly consistent setting.
\end{feedback}
See W2.

\paragraph{Reviewer 1: D4}
\begin{feedback}
  W3 would really help to showcase the value of segmentation in practice.
  Ideally, we want to see that in a specific application there is a logical
  mapping between segments and data, and then use this work as an example for
  potential future work that will be fully automated as the authors suggest.
\end{feedback}
See W3.

\paragraph{Reviewer 1: D5}
\begin{feedback}
  In fact, this automation would definitely strength the paper dramatically if
  it were added (but it is not required for this extended version to be enough)
\end{feedback}
We think that automated segmentation would be such a significant contribution
that it would warrant a publication on its own.

\paragraph{Reviewer 1: D6}
\begin{feedback}
  It is clear that the authors added more material to the journal paper,
  however, it does not substantially augment the messages of the paper --
  addressing W1-W3 will help in that direction (for example section 7 ends
  without any enough discussion of the practical connections of the proposed
  idea)
\end{feedback}
We disagree. Our paper is largely theoretical. We provide new, and arguably
improved, definitions of \invariantconfluence{}. We characterize the
relationship between \invariantclosure{} and \invariantconfluence{}. We exploit
the relationship to develop an \invariantclosure{} decision procedure. We
invent the notion of \mergereducibility{} and characterize its relationship
with \invariantclosure{}. We invent the notion of segmented
\invariantconfluence{}, and develop another decision procedure.

Addressing W1-W3 would certainly strengthen the paper. We don't argue that.
But, these three points focus exclusively on the empirical evaluation of our
techniques, rather than the theoretical underpinnings, which we feel is the
core message of the paper. This is why we focused our revisions on the
theoretical aspects of the paper. We believe that these additions are
substantial. Repeating from our change log:
\begin{itemize}
  \item (Section 2.3)
    In our conference paper, we present a definition of invariant confluence
    that is slightly different than the original definition from Bailis et
    al.~\cite{bailis2014coordination}, but we do not comment on the difference.
    We added a subsection (Section 2.3) clarifying that the difference in
    definitions is superficial. We describe Bailis' original definition using
    our formalism and prove that the two definitions are equivalent.

  \item (Section 3, Theorem 2)
    We added a formal proof that invariant closure is sufficient for invariant
    confluence.

  \item (Section 3, Theorem 3)
    We added a formal proof of our key result equating invariant closure and
    invariant confluence.

  \item (Section 4.1)
    We expanded the correctness proof of our interactive decision procedure. A
    key step in understanding the correctness of our interactive invariant
    confluence decision procedure---namely that $I$-reachable and
    $(I-NR)$-reachable states are equal---was omitted from our conference
    paper. We added a formal proof of this fact as well as a few other minor
    elaborations.

  \item (Section 4.2)
    In our conference paper, we asserted that checking for invariant closure is
    undecidable in general. We added a brief example showing \emph{why}
    invariant closure is undecidable, by reducing it to the Halting Problem.

  \item (Section 4.3)
    One of the core contributions of our paper is an interactive invariant
    confluence decision procedure. Here, ``interactive'' means that the
    decision procedure relies on human input to guide the decision procedure.
    It's possible for a person to provide the decision procedure with incorrect
    information. We added a subsection (Section 4.3) describing how to mitigate
    these errors.

  \item (Section 5, Examples 3 and 4)
    We added examples of merge-reducible and non-merge-reducible objects to
    make it easier for the reader to understand the definition of
    merge-reducibility.

  \item (Section 5, Theorem 4 and Figure 4)
    We added a formal proof that merge-reducibility is a sufficient condition
    for invariant-confluence. We also added a diagram illustrating the key idea
    behind the proof. Similarly, we added a counterexample showing that
    merge-reducibility is not a necessary condition for invariant-confluence.

  \item (Section 5, Theorem 5 and Figure 5)
    We added a formal proof that criteria 1 through 4 are sufficient for
    invariant confluence. The idea behind the proof is straightforward but the
    details are tedious. To help readers see the forest through the trees, we
    also added an illustration of the key idea of the proof.

  \item (Section 5, Example 5)
    In our conference paper, we provide an example of a distributed object that
    satisfies criteria 1 through 4 outlined in Theorem 4 but is not invariant
    closed. In the example, we asserted that criteria 1 through 4 are satisfied
    but did not elaborate on why. We added a brief explanation making it clear
    why the criteria are met.

  \item (Section 6.3, Example 7)
    We added an example of co-reachable and co-unreachable states to help the
    reader understand the definition of co-reachability.

  \item (Section 6.4)
    We added a paragraph explaining the relationship between segmented
    invariant confluence and a distributed locking approach to maintaining
    invariants (that can be found in related works).

  \item (Section 6.4)
    We added a paragraph explaining a couple of optimizations that can be made
    to our segmented invariant confluence system model to reduce the frequency
    of global coordination.

  \item (Section 6.4)
    We added a paragraph explaining how a segmented invariant confluence
    decision procedure can leverage our merge reducibility results in addition
    to our results on invariant closure.

  \item (Section 6.4)
    We clarified the fault tolerance limitations of segmented invariant
    confluence. A naive implementation of our segmented invariant confluence
    decision procedure does not guarantee liveness, but there are
    straightforward techniques we can leverage to introduce liveness. We added
    an explanation of this to the paper.

  \item (Section 7)
    Our conference paper presents formal definitions of distributed objects and
    invariant confluence. The definitions revolve around servers gossiping
    state to one another and merging the states together using a binary merge
    operator. This approach is called a ``state-based'' approach. We added a
    section (Section 7) outlining a dual ``operation-based'' approach in which
    transactions, rather than states, are gossiped. This duality between
    state-based and operation-based definitions is well established in the CRDT
    literature.
\end{itemize}

\paragraph{Reviewer 1: D7}
\begin{feedback}
  In the intro we read about ``the trade-off between application complexity and
  coordination-freedom'', it would be great to see some practical examples later
  in the paper
\end{feedback}
\textcolor{red}{TODO.}

### 2020-08-08 doc/the_vldb_journal_paper/sections/evaluation.tex
More revisions.
Only four left!

\section{Evaluation}\seclabel{Evaluation}
In this section, we describe and evaluate Lucy: a prototype implementation of
our decision procedures and system models.
\begin{revisions}
  All code is available online:
  \urlstyle{rm}
  \url{https://github.com/mwhittaker/enforced_invariant_confluence}.
\end{revisions}

\subsection{Implementation}
Lucy includes an implementation of the interactive decision procedure described
in \algoref{InteractiveDecisionProcedure}, an implementation of a decision
procedure that checks criteria (1) - (4) from \thmref{LatticeProperty}, and an
implementation of the decision procedure described in
\algoref{ArbitraryStartInteractiveDecisionProcedure}. The decision procedures
are implemented in roughly 2,500 lines of Python. Programmers specify objects,
transactions, and invariants in a small Python DSL and interact with the
interactive decision procedures using an interactive Python console. Note that
a programmer only has to run the decision procedures offline a single time to
check the \invariantconfluence{} of their distributed object. The decision
procedures do not have to be run online when transactions are being processed.

@@

  Escrow transactions are a good example where this optimization works
  particularly well. Assume that $k = 3$ and that the three states are $s_1 =
  s_2 = s_3 = (p_1, p_2, p_3), (n_1, n_2, n_3) = (3, 3, 3), (3, 3, 3)$. Here,
  no server can process a decrement transaction locally. If a server receives a
  decrement transaction, it must execute it with global coordination. However,
  if the server instead defers the execution of the decrement transaction and
  happens to receive a number of increment transactions, then it can execute
  the decrement locally, without the need for global coordination.
\end{revisions}

>>>>

\section{Evaluation}\seclabel{Evaluation}
In this section, we describe and evaluate Lucy: a prototype implementation of
our decision procedures and system models.
\begin{revisions}
  All code and all the specifications described below are available online:
  \urlstyle{rm}
  \url{https://github.com/mwhittaker/enforced_invariant_confluence}. The
  repository also contains a couple more example specifications.
\end{revisions}

\subsection{Implementation}
Lucy includes an implementation of the interactive decision procedure described
in \algoref{InteractiveDecisionProcedure}, an implementation of a decision
procedure that checks criteria (1) - (4) from \thmref{LatticeProperty}, and an
implementation of the decision procedure described in
\algoref{ArbitraryStartInteractiveDecisionProcedure}. The decision procedures
are implemented in roughly 2,500 lines of Python. Programmers specify objects,
transactions, and invariants in a small Python DSL and interact with the
interactive decision procedures using an interactive Python console.
\begin{revisions}
  The Python DSL includes a library of CRDTs including integers, booleans,
  tuples, sets, option types, and maps. It supports invariants involving
  equalities, inequalities, arithmetic, boolean logic, and basic set operations
  (e.g, union, intersection).
\end{revisions}\
Note that a programmer only has to run the decision procedures offline a single
time to check the \invariantconfluence{} of their distributed object. The
decision procedures do not have to be run online when transactions are being
processed.

@@

  Escrow transactions are a good example where this optimization works
  particularly well. Assume that $k = 3$ and that the three states are $(3, 3,
  3), (3, 3, 3)$. Here, no server can process a decrement transaction locally.
  If a server receives a decrement transaction, it must execute it with global
  coordination. However, if the server instead defers the execution of the
  decrement transaction and happens to receive a number of increment
  transactions, then it can execute the decrement locally, without the need for
  global coordination.
\end{revisions}
